Distribution-oblivious Programming with Dish

### 2019-11-20 pldi/draft.tex
Add a very drafty draft of the frond end section

@@

\section{Extensibility}

@@

\section{Intermediate Representation/Dataflow graph}

@@

\kk{Maybe this becomes a subsection either in the intermediate
  representation or in the planner.}

We parse a shell script using a posix compliant parser that returns
the ast.

\kk{Text point: The AST Is not a very good data structure to
  manipulate for distribution, so we translate it to a kind of a
  dataflow graph.}

@@

\kk{Text point: Problem: Not all parts of a shell script are
  distributable, so we cannot arbitrarily distribute. Here we have to
  make our assumptions clear}

Even though commands are split in categories, and several pipelines
are distributable, not all parts of a shell script are stateless and
several of them have effects that could affect the behaviour of a
program after distribution.

\kk{Solution: We have found that distributing subtrees of the shell
  AST is safe. Can we have a semi-formal argument about that?}

Mention all the different constructs that our front end handles, and
why it is safe to add them to the distribution pipeline.

\kk{The same file exists once in the substree. This is a conservative
  assumption that ensures that no two commands write or read to the
  same file concurrently.}

\kk{How can we make sure that nodes in the IR are deterministic. Do we
  make some assumption that enforces this?}

@@

Future work on this: calling the shell shtepper to only partially
evaluate strings, just expand arguments, and not do any significant
computation.

@@

\begin{itemize}
\item Fault Tolerance
\item Distributed Planning: The planner at the moment is naive, and
  there is a huge literature on operator placement of distributed
  dataflow graphs that can be used to improve it.
\item Support of shell constructs: At the moment we have been
  conservative in the shell subsets that we handle, so that we don't
  introduce any unsafe parallelism. The point of this work was not to
  be able to handle a complete distributable subset of the shell, but
  rather a significantly big part of it, to demonstrate that there can
  be performance benefit from it.
\item Support distributing different classes: At the moment we are
  only dealing with stateless and pure commands, that is the ones that
  are most popular but also can offer the biggest benefits from
  parallelizing them. However, it would be interesting to explore the
  distributability of the file system commands, by using a distributed
  file system. THis would also require considerations about
  replication, consistency, and failures.
\item No cycles
\end{itemize}

>>>>

\kk{The title is too abstract and doesn't really hint that we
  distribute shell code. Also we could consider capitalizing the name
  as DiSh?}

@@

\section{Extensibility}
\label{ext}

@@

\section{Intermediate Representation/Dataflow graph}
\label{ir}

@@

\kk{It might make sense to merge this section with the one below with
  the name: From Shell to IR and back}

As mentioned in \Cref{ir}, in order to generate the distributed
implementation of a shell script, \texttt{Dish} first translates it to
a dataflow graph. However, there are several components in shell
programs that cannot be arbitrarily parallelized without affecting the
program behaviour.

\kk{Give examples of writing to environment variables, producing
  irreversible side effects, the ; and \&\& operators.}

In order to address this issue, we introduce the notion of
distributable AST subtrees, and we design an analysis that given the
shell script Abstract Syntax Tree identifies distributable AST
subtrees and translates each one of them to a dataflow graph. The AST
of the shell script is produced using a POSIX compliant parser
\kk{Reference libdash github}.

@@

\kk{The goal of this section is to convince the reader ``intuitively''
  that we choose safe parts of a shell program to distribute. There is
  no formal argument, but one should be able to intuitively believe
  us.}

Consider the following example:

\kk{Give an example of a shell script that cannot be distributed
  completely, Possibly involving ; and writing to a variable.}

\kk{Intuitively explain what could be distributed and what not.}

\kk{Our notion of distributable subtrees has two dimensions. First,
  the state of the shell should be an invariant in a distributable
  subtree. We cannot know this completely without evaluating, so we do
  a conservative analysis. The second dimension is that there exists
  at least one component that can be abstracted as a node of the
  dataflow graph. This can be anything that includes a command.}

%% Even though commands are split in categories, and several pipelines
%% are distributable, not all parts of a shell script are stateless and
%% several of them have effects that could affect the behaviour of a
%% program after distribution.

\kk{All of our assumptions and study are based on the POSIX
  standard. Reference?}

\kk{Not checked Assumption. A command doesn't write/read to arbitrary
  files that are not mentioned in its arguments. This is necessary to
  enforce that no two commands concurrently read/write from the same
  file.}

\kk{Not checked Assumption: No signals, traps, and sets at the
  moment. At the moment we don't have a way to handle signals and
  traps.}

\kk{Partially checked Assumption. The same file exists once in the
  substree. This is a conservative assumption that ensures that no two
  commands write or read to the same file concurrently. We kind of
  check it but not completely. We check that the expanded evaluated
  files are not the same, but one could have part of a file in a
  variable etc.}

\kk{Checked assumption: How can we make sure that nodes in the IR are
  deterministic. Do we make some assumption that enforces this? I
  guess that commands that are nodes in the IR are all deterministic?
  That is stateless and pure commands are deterministic.}

\paragraph{analysis}

The analysis is done in a DFS fashion on the AST. Distributable
subtrees grow from the bottom up, until a construct that doesn't
``allow'' them to go through is encountered.

%% Mention all the different constructs that our front end handles, and
%% intuitively mention why it is ``safe'' to add them to the distribution
%% pipeline.

Subtrees don't cross the following:

\begin{itemize}
\item Sequence, And,  or
\item Function definitions
\item Command Substitutions
\item control flow: while, for, if, case: While these constructs can
  be parallelized and there has been research on it, it is orthogonal
  to our work, and can be incorporated as future work.
\end{itemize}

On the other hand: A distributable subtree starts whenever there is a
pipe, or asynchronous list. If the subcomponents of the pipe or
asynchronous list were already graphs, then we connect them to the
bigger graph \kk{explain how we do pipe and asynchronous list}, and if
they were not graphs, we make each subcomponent a node in the subtree,
event though they are not commands. These nodes do not belong to any
category, but still have stdin and stdout, and can be executed
concurrently (as long as they satisfy the checked assumptions above),
since they don't introduce new behaviours, they would still be
executed concurrently when ran in the standard shell, producing
possibly non-deterministic results with different interleavings.

\kk{Explain how we merge two graphs when seeing a pipe}

Whenever we see a subtree of the AST that is a command, we check the
files that it writes to. If any of its arguments is a command
substitution or contains anything that is evaluated\kk{make this more
  precise}, its category is marked as sequential\kk{name?}. Otherwise
we check its its category and its input output arguments (using the
methods described in \Cref{ext}). The command is then initialized as
the node of a singular graph, with the inputs and outputs of the graph
being the ones of the node.

\kk{Maybe mention the recursive case. I am not sure if that should be
  here, or in the distribution planner. In any case, it should be
  after we say that an IR is replaced with a call to the planner with
  the IR saved in a file. The example is a command in a pipeline whose
  first argument is a command substitution with a pipeline. Executing
  the outside dataflow graph, would lead to the command to be executed
  as part of this shell. However its first argument would have a call
  to the planner itself, leading to another distributed implementation
  running in parallel. The rest of the pipeline would wait for this to
  end before it continues. Assuming that the files that both of these
  graphs use are different, and that the above assumptions hold, there
  should be no interference between the two graphs and they should be
  able to run in parallel}

@@

Future work on this: calling the shell shtepper to only partially
evaluate strings, just expand arguments, and not do any significant
computation. Even better, we could be calling the shtepper just after
having parsed the ast, and decide on maximal distributable subtrees as
late as possible in the process. This would be awesome.

\section{Implementation}

\kk{We should make sure to measure the time that it takes for the
  whole process to run (parsing, translating, optimizing, planning)
  together with the scripts. This will probably be negligible, but we
  still have to mention it. }

\kk{Also, we have to have one-two handcrafted examples that have more
  than one pipeline, to show the execution time end to end, as this
  cannot be shown with the one-pipeline examples since they only have
  one graph, so no back and forth between shell and dish.}

\kk{Furthermore it would be great to have one distributed example that
  runs in more than one node (if we have time)}

\kk{Also we have to make sure to measure with and without the final
  cat of output file, so that we show that this is also negligible.}

\kk{We have to make sure that we repeat that the results are the same
  in all tests that we run. Corectness, blah blah ...}

@@

\begin{itemize}
\item Fault Tolerance
\item Distributed Planning: The planner at the moment is naive, and
  there is a huge literature on operator placement of distributed
  dataflow graphs that can be used to improve it.
\item Support of shell constructs: At the moment we have been
  conservative in the shell subsets that we handle, so that we don't
  introduce any unsafe parallelism. The point of this work was not to
  be able to handle a complete distributable subset of the shell, but
  rather a significantly big part of it, to demonstrate that there can
  be performance benefit from it.
\item Support distributing different classes: At the moment we are
  only dealing with stateless and pure commands, that is the ones that
  are most popular but also can offer the biggest benefits from
  parallelizing them. However, it would be interesting to explore the
  distributability of the file system commands, by using a distributed
  file system. THis would also require considerations about
  replication, consistency, and failures.
\item No cycles (multiple commands writing and reading from the same file)
\item Formal proof that the analysis indeed returns distributable
  subtrees. Previously there was no formal semantics for shell, but
  recent work (cite greenberg) allows us to make a formal proof that
  the analysis is indeed sound.
\end{itemize}
